---
title: "Week 9_Raster Interpolation"
author: "Wanjiku Ngare"
date: "11/2/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(sf)
library(leaflet)
library(htmltools)
library(htmlwidgets)
library(raster)
library(gstat)
library(spatial)
```

```{r, include=FALSE}
boston_nhoods <- st_read("https://opendata.arcgis.com/datasets/3525b0ee6e6b427f9aab5d0a1d0a1a28_0.geojson", quiet = TRUE) %>%
  dplyr::select(Name)

leaflet(boston_nhoods) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(highlightOptions = highlightOptions(fillColor = "yellow", 
                                                  fillOpacity = 1),
              label = ~Name, 
              weight = 1) 

```

## Total number of people tested for coronavirus by neigbhorhood in Boston

The data about number of people tested for COVID by neighborhood came from the Boston Public Health Commission. While it was subdivided by neighborhood, it was missing some neighborhoods and combined others. There was no data for Longwood and Harbor Island - likely because they don't have many residents. Data for West Roxbury was included in the data for Roslindale. Brighton was included in Allston; Beacon Hill, West End, North End, Bay Village, and Downtown (which includes Chinatown and Leather District) were included in Backbay; and South Boston Waterfront in South Boston. Where the number tested is indicated as 1, the neighborhood's data was merged with an adjacent neighborhood described above. 

```{r, echo=FALSE}
boston_nhoods <- boston_nhoods %>%
  mutate(number_tested = c(24806, 22693, 1, 1, 1, 1, 1, 1, 27649, 23352, 31671, 23354, 8492, 1, 1, 1, 48991, 1, 1, 14681, 9933, 57347, 1, 20109, 37121, 1))

boston_nhoods$label <- 
  paste(boston_nhoods$Name, "<br>", 
        boston_nhoods$number_tested, " Total tests") %>% 
  lapply(htmltools::HTML)

bins <- seq(min(boston_nhoods$number_tested),
            max(boston_nhoods$number_tested), by = 5000)
pal <- colorNumeric("viridis", 
                    domain = boston_nhoods$number_tested,
                    na.color = "#00000000")

leaflet(boston_nhoods) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(highlightOptions = highlightOptions(fillOpacity = 1),
              label = ~label,
              fillColor = ~pal(number_tested),
              weight = 1, color = "black") %>% 
  addLegend(pal = pal, 
            values = ~number_tested,
            bins = 5,
            opacity = 0.7, title = "Total tests by neighborhood",
            position = "topright")

```

Total number of people tested for coronavirus by neigbhorhood in Boston as point data

```{r, include=FALSE}
MA_state_plane <- "+proj=lcc +lat_1=41.71666666666667 +lat_2=42.68333333333333 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs "

WGS84 <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"

nhood_points <- st_centroid(
  st_transform(boston_nhoods, crs = MA_state_plane)) %>%
  st_transform(WGS84)

leaflet(nhood_points) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addCircles(label = ~label,
             fillColor = ~pal(number_tested),
             stroke = FALSE, 
             radius = 50, 
             fillOpacity = 1) %>% 
  addLegend(pal = pal, 
            values = ~number_tested,
            bins = 5,
            opacity = 0.7, title = "Total number of people tested by neighborhood",
            position = "topright")
```


```{r, include=FALSE}
nhood_pts_sp <- nhood_points %>%
  st_transform(MA_state_plane) %>%
  as_Spatial()

nhood_poly_sp <- boston_nhoods %>%
  st_transform(MA_state_plane) %>%
  as_Spatial()
```

```{r, include=FALSE}
boston_raster <- raster(nhood_poly_sp, res=10)
```

```{r, include=FALSE}
gs <- gstat(formula=number_tested~1, locations=nhood_pts_sp)
idw_interp <- interpolate(boston_raster, gs)

idw_interp_clip <- mask(idw_interp, nhood_poly_sp)
```

# Interpolation map of total number of coronavirus test administered in Boston neighborhoods

This interpolation is both helpful in filling in information regarding what is happening between neighborhoods, but also not completely accurate because I've input dummy data of "1" for neighborhoods that were aggregated by the Boston Public Health Commission, as described in the first map. 
```{r, echo=FALSE,warning=FALSE}
leaflet(nhood_points) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addRasterImage(idw_interp_clip, colors = pal, opacity = 0.8) %>% 
  addLegend(pal = pal, 
            values = ~number_tested,
            bins = 5,
            opacity = 0.7, title = "Number of covid tests<br>administered by neighborhood",
            position = "topright")
```























